\chapter{Acknowledgements}

I would like to thank my wife Gazala who supported me and pushed me to do more astronomy and apply to the Master's program at Macquarie University. I would also like to thank my family in Sri Lanka who support everything I do and encouraged me to follow my interests no matter what. I would not be completing this work if not for my family.

A big thank you to my supervisor Dr Daniel Zucker who in addition to providing valuable feedback, guided and believed in me every step of the way. Your friendship and mentorship has made this journey so much more enjoyable. Thank you to Dr Ben Montet at UNSW who introduced me to the astronomy community in Australia and encouraged me to apply to Macquarie University. As an immigrant and someone who had no formal training in astronomy until this point, I owe a debt of gratitude.

Many thanks to Dr Sarah Martell at UNSW for the early feedback and guidance on P Cygni spectra. Thank you to Dr Gregor Traven at Lund University for his valuable feedback and sanity checks towards the end of the project. Thanks to Arv Hughes, Dr Sven Buder and Dr Klemen Čotar and the various members of the GALAH science team and HDR cohort at Macquarie University who provided feedback and support during various forums and meetings throughout the last year - your kindness, generosity and camaraderie is highly appreciated. Finally, many thanks to all my friends, particularly Nuzhi Meyen in Sri Lanka for his input on the mathematics of machine learning and my many friends in Sydney who have had to endure me talking about astronomy 24/7/365.